\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-01-30 forwards-and-backwards.tex
edits!

% % % % % % % % % % % % %
\subsection{Tightness of the Lines of Descent}
    \label{sec:lookdown_convergence}

@@

The process we consider is similar to the state-dependent branching processes
of \citet{KR}, so one might expect that the proofs there would carry over with little change.
However, there is an important difference:
Recall that the level $U_a(t)$ of a line of descent evolves as
$$
\dot u 
=
c(x, \eta)
u^2
- b(x, \eta)
u,
$$
where
\begin{align*}
    c(x, \eta) &= \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) \\
    \text{and}\qquad
    b(x, \eta)
    &=
    \theta \gamma(x, \eta) \int_{\IR^d} \left(r(y, \eta) - r(x, \eta)\right) q(x, dy)
    + F(x, \eta) 
\end{align*}
(see equations \eqref{eqn:level_generator} and \eqref{eqn:b_limit}),
where $c(x, \eta) \ge 0$ and $b(x, \eta)$ may take either sign.
In \citet{kurtz/rodrigues:2011}, $b$ was bounded above and $c$ was bounded away from zero,
so they noted that if $U_a(t) \ge b/c$
lines of descent would only move upwards from that time onwards.
This allows the processes to be jointly and simultaneously constructed
for all values of $N$, with a pointwise embedding
of $(\lp^N_t)_{t \ge 0}$ within $(\lp^{M})_{t \ge 0}$ for $b/c < N < M$.
In other words, individuals with levels above $N > b/c$ at time $t_0$
do not affect $(\lp^N_t)_{t \ge t_0}$,
thus allowing a comparison of the number of lines of descent below level $u_0$
to a branching process.
Although we have provided a
joint construction of $\lp^N$ for all $N$ in Section~\ref{sec: individual lines of descent},
it does not have this monotonicity:
for one thing, $b$ and $c$ depend on the population process $\eta$
and so all individuals can affect all other ones (even those with lower levels).
Furthermore, in the deterministic case $\theta/N$ and hence $c$ converge to zero,
and so lines of descent with arbitrarily high level may drift back downwards.
Indeed, this must be the case if the population persists,
since in the deterministic case there is no branching.
In our case $b$ and $c$ depend on the state of the population,
so we have no such monotonicity.

@@

Let $C_c^+$ denote the set of nonnegative continuous functions on $\overline{\IR^d} \times [0, \infty)$
for which there exists a $u_f \ge 0$ such that $u > u_f$ implies that $f(x, u) = 0$.
Recall that we are working with the vague topology ``in the level direction'',
so that $m_n \to m$ in $\lpmeasures$ if and only if $\langle f, m_n \rangle \to \langle f, m \rangle$
for all $f \in C_c^+$.
By Theorem~A.23 in \citet{kallenberg1997foundations},
the space $\lpmeasures$ is separable and metrizable,
and $K \subset \lpmeasures$ is relatively compact if and only if
$\sup_{\lp \in K} \langle f, \lp \rangle < \infty$
for any $f \in C_c^+$.
% The compact sets we consider are those with bounded total mass of their spatial projection,
% i.e., of the form
% $$
%     K_M = \left\{
%         \lp \in \lpmeasures \;:\;
%         \sup_{L \ge 1} \frac{1}{L} \lp\left(\overline{\IR^d} \times [0, L]\right) \le M
%     \right\} .
% $$
% These sets are compact 
% because if $\lp \in K_M$ and $f \in C_c^+$ then $\langle f, \lp\rangle \le M u_f \|f\|_\infty$.
Let $\mathcal{H} = \{h_n\}_{n \ge 1}$ be a sequence of
monotone, nonnegative, compactly supported, differentiable functions
on $[0, \infty)$ such that there exist sequences $u_n \to \infty$ and $c_n$ such that
\begin{enumerate}
    \item $h_n(u) > 1$ for $u \le u_n/2$,
    \item $h_n(u) = 0$ for $u \ge u_n$,
    \item $- h_n(u) \le h_n'(u) \le 0$, and
    \item $\max\left( \int_u^\infty h_n(v) dv, \int_u^\infty h_n(v)^2 dv \right) \le c_n h_n(u)$.
\end{enumerate}
For instance, we could take $h_n(u)$ to be a mollified version of $\min(1, e^{-(u-n)})$,
and then $u_n = c_n = n$.
The compact sets we consider are those with uniformly bounded integrals against $\mathcal{H}$,
i.e., of the form
$$
    K_C = \left\{
        \lp \in \lpmeasures \;:\;
        \langle h_n, \lp \rangle \le C_n
        \text{ for all } n \ge 1
    \right\} ,
$$
for some sequence of constants $(C_1, C_2, \ldots)$.
This set is compact because any $f \in C_c^+$ is bounded by a constant multiple of some $h_n$.
\comment{check: why is it closed?}

\begin{lemma}[Compact containment for $\lp$]
    \label{lem:compact containment lookdown}
    Under Assumptions~\ref{},
    if $\theta/N \to \alpha$ and $\lp_0^N \to \lp_0$ weakly as $N \to \infty$
    \comment{and they are conditionally Poisson},
    for any $T$ and $\delta > 0$ there exists a sequence
    $(C_1, C_2, \ldots)$ of finite constants such that
    \begin{align} \label{eqn:compact eqn lookdown}
    \limsup_{N \to \infty}
    \IP\left\{
        \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > C_n
        \text{ for some } n \ge 1
    \right\}
    < \delta .
    \end{align}
\end{lemma}

@@

\begin{proof}[of Lemma~\ref{lem:compact containment lookdown}]
    By a union bound,
    $$ 
    \IP\left\{
        \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > C_n
        \text{ for some } n \ge 1
    \right\}
    \le
    \sum_{n \ge 1}
    \IP\left\{
        \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > C_n
    \right\} ,
    $$
    so \eqref{eqn:compact eqn lookdown} follows
    by taking $C_n = 2^{n-1} K_n / \delta$ in Lemma~\ref{lem:compact containment hn}.
\end{proof}

\begin{proof}[of Lemma~\ref{lem:compact containment hn}]
    \comment{TODO: just give bounds on moments and say this follows
        the proof of Lemma~\ref{lem:eta_compact_containment} exactly, otherwise.
        Or make the basic idea a lemma in the appendix?}
    Let $h \in \mathcal{H}$.
    It is evident from the description of the process
    (or, by differentiating in Definition~\ref{defn:lookdown_mgale})
    that
    \begin{align*}
    \begin{split}
        \langle h^n(u), \lp_t^N(dx, du) \rangle
        &=
        \langle h^n(u), \lp_0^N(dx, du) \rangle
        + M^{h_n}_t 
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            2 a^N(x, \eta^N_s) \int_u^N h_n(u_1) du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            h_n'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
    \end{split}
    \end{align*}
    where $M^{h_n}_t$ is a martingale
    with quadratic variation
    \begin{align*}
        [M^{h_n}]_t
        &=
        \int_0^t
        \langle 2 a^N(x, \eta_s^N) \int_u^N h(u_1)^2 du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}
    Now, note that $0 \le a^N(x, \eta_x^N) \le C_a$ for some $C_a < \infty$,
    and we have assumed that $\int_u^\infty h_n(u_1) du_1 \le C_n h_n(u)$,
    so that 
    \begin{align*}
        & \int_0^t
        \big\langle
            2 a^N(x, \eta^N_s) \int_u^N h(u_1) du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad
        \le
        2 C_a C_n \int_0^t \big\langle h_n(u), \lp_s^N(dx, du) \big \rangle .
    \end{align*}
    In the next line, the contribution from $a^N(x, \eta^N_s) u^2 h'(u)$ is nonpositive because
    $a^N \ge 0$ and $h'(u) \le 0$.
    Similarly, since $b^N(x, \eta_x^N) \le C_b$ for some $C_b < \infty$,
    and by assumption
    $- h'(u) \le C_n h(u)$ and $h'(u) = 0$ for $u \ge u_n$,
    and so
    \begin{align*}
        & \int_0^t
        \big\langle
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            h_n'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \le
        C_b u_n C_n \int_0^t \big\langle h_n(u), \lp_s^N(dx, du) \big \rangle .
    \end{align*}
    Putting this together, we have that
    \begin{align} \label{eqn:hn_bound}
        \langle h^n, \lp_t^N \rangle
        &\le
        \langle h^n, \lp_0^N \rangle
        + M^{h_n}_t 
        + (2 C_a + C_b u_n) C_n \int_0^t 
            \langle h^n, \lp_s^N \rangle
        ds .
    \end{align}
    By Gronwall's inequality, with $C = (2 C_a + C_b u_n) C_n$
    \begin{align*}
        \IE\left[ \langle h^n, \lp_t^N \rangle \right]
        &\le
        \IE\left[ \langle h^n, \lp_0^N \rangle \right] e^{Ct} .
    \end{align*}

    Now consider the maximum. By \eqref{eqn:hn_bound},
    \begin{align*}
        \sup_{0 \le t \le T} \langle h^n, \lp_t^N \rangle
        &\le
        \langle h^n, \lp_0^N \rangle
        + \sup_{0 \le t \le T} M^{h_n}_t 
        + C \int_0^t 
            \sup_{0 \le t \le T} \langle h^n, \lp_s^N \rangle
        ds .
    \end{align*}
    Again since $\sqrt{x} \le 1 + x$ for $x \ge 0$,
    The Burkholder-Davis-Gundy inequality
    tells us that there is a $C'$ such that
    \begin{align*}
    \IE\left[ \sup_{0 \le t \le T} M^{h_n}_t \right]
    & \le
        C'\left( 1 + \IE\left[ [M^{h_n}]_t \right] \right)
    \\ & \le
        C' \left( 1 + \int_0^t \IE\left[
        \langle 2 a^N(x, \eta_s^N) \int_u^N h(u_1)^2 du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        \right] ds \right) 
    \\ & \le
        C' \left( 1 + 2 C_a C_n \int_0^t \IE\left[
        \langle h_n, \lp_s^N \big \rangle
        \right] ds \right) 
    \\ & \le 
        C' \left( 1 + C'' e^{Ct} \IE[ \langle h^n, \lp_t^N \rangle ] \right) ,
    \end{align*}
    where $C'' = 2 C_a C_n / C$.
    Therefore, 
    \begin{align*}
        \IE\left[ \sup_{0 \le t \le T} \langle h^n, \lp_t^N \rangle \right]
        &\le
        C' + \IE\left[ \langle h^n, \lp_0^N \rangle \right]( 1 + C'' e^{Ct} )
        + C \int_0^t 
            \IE\left[ \sup_{0 \le t \le T} \langle h^n, \lp_s^N \rangle \right]
        ds ,
    \end{align*}
    and so once again by Gronwall's inequality, there is a $C'''$ such that
    \begin{align*}
        \IE\left[ \sup_{0 \le t \le T} \langle h^n, \lp_t^N \rangle \right]
        &\le
        C'''(1 + \IE[\langle h^n, \lp_0^N \rangle]) e^{2Ct} ,
    \end{align*}
    and so
    \begin{align*}
        \IP\left\{ \sup_{0 \le t \le T} \langle h^n, \lp_t^N \rangle > K \right\}
        &\le
        \frac{
            C'''(1 + \IE[\langle h^n, \lp_0^N \rangle]) e^{2CT} ,
        }{ K } .
    \end{align*}
\end{proof}

We record for future reference
the Aldous-Rebolledo criteria for tightness of a sequence real-valued processes:
\comment{TODO: reference this elsewhere, if we keep it?}

@@

\begin{proof}[of Lemma \ref{lem:test_fn_tightness}]
    First, a reminder of the description of the process
    \comment{(TODO: figure out if the notation here for $a$ and $b$ should be used above;
        if so, move it up; if not, replace it here.)}
    Recall when the state of the process is $\eta$,
    the level of a line of descent at spatial location $x$ and level $u$
    changes following
    $$
        \dot u = a^N(x, \eta) u^2 - b^N(x, \eta) u,
    $$
    where
    $$
        a^N(x, \eta)
        =
        \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) ,
    $$
    and
    $$
        b^N(x, \eta)
        =
        \theta \gamma(x, \eta) \int_{\IR^d} (r(y, \eta) - r(x, \eta)) q(x, dy) + F(x, \eta) .
    $$
    Furthermore, the line of descent
    gives birth to lines at higher levels at rate $2 (N - u) a^N(x, \eta)$,
    that each such new line chooses a level uniformly from $[u, N]$,
    a spatial location $y$ from the kernel
    $$
        q^m(x, dy, \eta)
        =
        r(y, \eta) q(x, dy) / \int_{\IR^d} r(y, \eta) q(x, dy),
    $$
    and the two lines swap spatial locations with probability 1/2.

    It is evident from the description of the process
    (or, by differentiating in Definition~\ref{defn:lookdown_mgale})
    that
    \begin{align*}
    \begin{split}
    M^f_t
        &=
        \langle f, \lp_t^N \rangle
        -
        \langle f, \lp_0^N \rangle
    \\ & \qquad \qquad {}
        -
        \int_0^t
        \big\langle
            a^N(x, \eta^N_s)
            \int_u^N \int_{\IR^d}
            \left(
                f(y, u_1) + f(x, u_1) + f(y, u) - f(x, u)
            \right)
            du_1 q^m(x, dy, \eta^N_s)
    \\ & \qquad \qquad \qquad \qquad {}
            +
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            \frac{d}{du} f(x, u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
    \end{split}
    \end{align*}
    is a martingale,
    with quadratic variation
    \begin{align*}
        [M^f]_t
        &=
        \int_0^t
        \langle a^N(x, \eta_s^N) \int_u^N \int_{\IR^d}
            f(y, u_1)^2 +
            \left( f(x, u_1) + f(y, u) - f(x, u) \right)^2
        du_1 q^m(x, dy, \eta^N_s)
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}

    Now we would like to bound the terms in the integral,
    to arrive at a bound using Gronwall's inequality, as before.

    First, note that there exists a constant $C_a$
    such that $a^N(x, \eta) \le C_a$
    (since $\gamma$ and $r$ are bounded above, and $\theta/N \to \alpha \ge 0$),
    and for any $\epsilon > 0$ and $T$ there is a $C_b$ such that for all $N$,
    $$
        \IP\left\{
            \sup_{0 \le s \le T} \sup_x |b^N(x, \eta_s^N)| > C_b
        \right\}
        < \epsilon 
    $$
    \comment{reference where we show this above}.
    Now, since $f$ vanishes for $u \ge u_0$,
    on the event that $\sup_x |b^N(x, \eta_s^N)| \le C_b$ then
    if $N \ge u_0$,
    \begin{align*}
    &
        \int_{\IR^d} \int_0^N
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            \frac{d}{du} f(x, u)
        \lp_s^N(dx, du)
    \\  & \qquad
    \le
        \int_{\IR^d} \int_0^{u_0}
            \left(
                C_a u_0^2
                + C_b u_0
            \right)
            \left\| \frac{d}{du} f \right\|_\infty
        \lp_s^N(dx, du) .
    \\  & \qquad
    \le
        C_0 \lp_s^N(\IR^d \times [0, u_0]) .
    \end{align*}

    Next, observe that for any $\eta$,
    $$
    \int_{\IR^d} \int_u^N (f(y, u_1) + f(x, u_1)) du_1 q^m(x, dy, \eta)
    \le 2 u_0 \|f\|_\infty .
    $$
    Furthermore,
    recall that we have assumed \comment{(check where we do this)}
    that $q^m$ is uniformly elliptic,
    and so since $f$ has bounded first and second derivatives,
    there is a $C_1$ depending on $f$ such that for all $\eta$,
    $\theta \int_{\IR^d} (f(y, u) - f(x, u)) q^m(x, dy, \eta) \le C_1$.
    Therefore, using boundedness of $\theta/N$,
    \begin{align*}
        &
        \int_{\IR^d} \int_u^N (f(y, u) - f(x, u)) du_1 q^m(x, dy, \eta^N_s)
        \\ &\qquad
        = (N - u) \int_{\IR^d} (f(y, u) - f(x, u)) q^m(x, dy, \eta^N_s)
        \\ &\qquad
        = \frac{N}{\theta} \left(1 - \frac{u}{N}\right) \theta \int_{\IR^d}
            (f(y, u) - f(x, u)) q^m(x, dy, \eta^N_s)
        \\ &\qquad
        \le C_1 \frac{N}{\theta} \ind_{u \le u_0} .
    \end{align*}

    Putting this together,
    \begin{align*}
        \IE\left[
            \langle f, \lp_t^N \rangle
        \right]
        \le
    \end{align*}

\subsubsection{OLD STUFF REMOVE}

    To get around this,
    first note that both $b$ and $c$ are uniformly bounded above 
    by Lemma~\ref{lem:gamma_bound} and since $\gamma$, $r$, $F$ \comment{(???)}, and $\theta/N$ are.
    Also note that for any $\delta > 0$ we can find a $K$ such that
    $$
    \IP\left\{
        \sup_{0 \le t \le T} \sup_x \smooth{F} \eta_t(x) < K
        \right\} < \delta,
    $$
    uniformly in $N$ and over a compact set of initial conditions.
    Since $F$ is uniformly continuous, this implies that $\sup_{0 \le t \le T} \sup_x F(x, \eta_t)$
    is similarly tight,
    as well as the rate of downward drift of the level of a line of descent,
    i.e., there is a $\bar b$ such that $b(x, \eta_t) \le \bar b$
    on a set of probability at least $1-\delta$ as well.
    Therefore, if $u_1 \ge u_0 + \bar b T$, then
    $$
    \IP\left\{
        \inf_{0 \le t \le T} U_a(t) \le u_0
        \;\big|\;
        U_a(0) > u_1
    \right\}
    \le \delta ,
    $$
    and so with high probability any lines of descent that had level below $u_0$
    at some point $0 \le t \le T$
    are descended from lines of descent with initial level below $u_1$.
    Since the number of lines of descent with level below $u_1$
    converges to a Poisson with finite mean,
    we need only show that the number of offpring lines of descent
    produced by this initially finite number is also tight.
    However, this follows immediately
    since the rate of production of new lines of descent with level below $u_1$
    of a line of descent at location $x$ and level $u$ is
    $2 (u_1 - u) N^{-1} \theta \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy)$
    and therefore uniformly bounded (because we may take $u_1 \le N$).

>>>>

% % % % % % % % % % % % %
\subsection{Tightness of the Lookdown Process}
    \label{sec:lookdown_convergence}

@@

The process we consider is similar to the state-dependent branching processes
of \citet{KR}, so one might expect that the proofs there would carry over with little change.
However, there is an important difference:
Recall that the level $U_a(t)$ of a line of descent evolves as
\comment{TODO: harmonize notation with above}
$$
\dot u 
=
c^N(x, \eta)
u^2
- b^N(x, \eta)
u,
$$
where
\begin{align*}
    c^N(x, \eta) &= \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) \\
    \text{and}\qquad
    b^N(x, \eta)
    &=
    \theta \gamma(x, \eta) \int_{\IR^d} \left(r(y, \eta) - r(x, \eta)\right) q(x, dy)
    + F(x, \eta) 
\end{align*}
(see equations \eqref{eqn:level_generator} and \eqref{eqn:b_limit}), where
$c^N(x, \eta) \ge 0$ and $b^N(x, \eta)$ may take either sign.
Assumptions~\ref{def:model_setup} imply that $c^N(x, \eta)$ is bounded
and $b^N(x, \eta)$, because of $F(x, \eta)$, is bounded above but not necessarily below.
In \citet{kurtz/rodrigues:2011}, $b^N$ was bounded above and $c^N$ was bounded away from zero,
so they noted that if $U_a(t) \ge b^N/c^N$
lines of descent would only move upwards from that time onwards.
This allows the processes to be jointly and simultaneously constructed
for all values of $N$, with a pointwise embedding
of $(\lp^N_t)_{t \ge 0}$ within $(\lp^{M})_{t \ge 0}$ for $b^N/c^N < N < M$.
In other words, individuals with levels above $N > b^N/c^N$ at time $t_0$
do not affect $(\lp^N_t)_{t \ge t_0}$,
thus allowing a comparison of the number of lines of descent below level $u_0$
to a branching process.
Although we have provided a
joint construction of $\lp^N$ for all $N$ in Section~\ref{sec: individual lines of descent},
it does not have this monotonicity:
for one thing, $b^N$ and $c^N$ depend on the population process $\eta$
and so all individuals can affect all other ones (even those with lower levels).
Furthermore, in the deterministic case $\theta/N$ and hence $c$ converge to zero,
and so lines of descent with arbitrarily high level may drift back downwards.
Indeed, this must be the case if the population persists,
since in the deterministic case there is no branching.
In our case $b^N$ and $c^N$ depend on the state of the population,
so we have no such monotonicity.

To complete the description of the process,
each line of descent
gives birth to lines at higher levels at rate $2 (N - u) c^N(x, \eta)$,
and each such new line chooses a level uniformly from $[u, N]$,
a spatial location $y$ from the kernel
$$
    q^m(x, dy, \eta)
    =
    r(y, \eta) q(x, dy) / \int_{\IR^d} r(y, \eta) q(x, dy),
$$
and the two lines swap spatial locations with probability 1/2.

It is evident from the description of the process
(or, by differentiating in Definition~\ref{defn:lookdown_mgale})
that
\begin{align} \label{eqn:f_xi_mgale}
\begin{split}
    \langle f, \lp_t^N \rangle
    &=
    \langle f, \lp_0^N \rangle
    + M^f_t
\\ & \qquad {}
    +
    \int_0^t
    \big\langle
        c^N(x, \eta^N_s)
        \int_u^N \int_{\IR^d}
        \left(
            f(y, u_1) + f(x, u_1) + f(y, u) - f(x, u)
        \right)
        du_1 q^m(x, dy, \eta^N_s)
\\ & \qquad \qquad \qquad \qquad {}
        +
        \left(
            c^N(x, \eta^N_s) u^2
            - b^N(x, \eta^N_s) u
        \right)
        \frac{d}{du} f(x, u)
    ,
    \lp_s^N(dx, du) \bigg \rangle
    ds ,
\end{split}
\end{align}
where $M^f$ is a martingale
with quadratic variation
\begin{align} \label{eqn:f_xi_qv}
    [M^f]_t
    &=
    \int_0^t
    \bigg \langle c^N(x, \eta_s^N) \int_u^N \int_{\IR^d}
        \big[ f(y, u_1)^2
\\&\qquad \qquad {}
        + \left( f(x, u_1) + f(y, u) - f(x, u) \right)^2 \big]
    du_1 q^m(x, dy, \eta^N_s)
    ,
    \lp_s^N(dx, du) \bigg \rangle
    ds .
\end{align}

In addition to tightness of the measure-values processes $\lp^N$,
the bounds used in the proofs below also imply
tightness of the number of lines of descent and the number of births below a fixed level,
and of the motion of individual lines of descent.
In other words, the limiting ``line of descent'' construction holds.

@@

Let $C_c^+$ denote the set of nonnegative continuous functions on $\overline{\IR^d} \times [0, \infty)$
for which there exists a $u_f \ge 0$ such that $u > u_f$ implies that $f(x, u) = 0$.
Recall that we are working with the vague topology ``in the level direction'',
so that $m_n \to m$ in $\lpmeasures$ if and only if $\langle f, m_n \rangle \to \langle f, m \rangle$
for all $f \in C_c^+$.
By Theorem~A.23 in \citet{kallenberg1997foundations},
the space $\lpmeasures$ is separable and metrizable,
and $K \subset \lpmeasures$ is relatively compact if and only if
$\sup_{\lp \in K} \langle f, \lp \rangle < \infty$
for any $f \in C_c^+$.
% The compact sets we consider are those with bounded total mass of their spatial projection,
% i.e., of the form
% $$
%     K_M = \left\{
%         \lp \in \lpmeasures \;:\;
%         \sup_{L \ge 1} \frac{1}{L} \lp\left(\overline{\IR^d} \times [0, L]\right) \le M
%     \right\} .
% $$
% These sets are compact 
% because if $\lp \in K_M$ and $f \in C_c^+$ then $\langle f, \lp\rangle \le M u_f \|f\|_\infty$.
Let $\mathcal{H} = \{h_n\}_{n \ge 1}$ be a sequence of
monotone, nonnegative, compactly supported, differentiable functions
on $[0, \infty)$ such that there exist sequences $u_n \to \infty$ and $k_n$ such that
\begin{enumerate}
    \item $h_n(u) > 1$ for $u \le u_n/2$,
    \item $h_n(u) = 0$ for $u \ge u_n$,
    \item $- h_n(u) \le h_n'(u) \le 0$, and
    \item $\max\left( \int_u^\infty h_n(v) dv, \int_u^\infty h_n(v)^2 dv \right) \le k_n h_n(u)$.
\end{enumerate}
For instance, we could take $h_n(u)$ to be a mollified version of $\min(1, e^{-(u-n)})$,
and then $u_n = k_n = n$.
\comment{(or we could just take those functions? they are differentiable almost everywhere.)}
The compact sets we consider are those with uniformly bounded integrals against $\mathcal{H}$,
i.e., of the form
$$
    K_C = \left\{
        \lp \in \lpmeasures \;:\;
        \langle h_n, \lp \rangle \le C_n
        \text{ for all } n \ge 1
    \right\} ,
$$
for some sequence of constants $(C_1, C_2, \ldots)$.
This set is compact because any $f \in C_c^+$ is bounded by a constant multiple of some $h_n$.

\begin{lemma}[Compact containment for $\lp$]
    \label{lem:compact containment lookdown}
    Under Assumptions~\ref{def:model_setup},
    if $\theta/N \to \alpha$ and $\lp_0^N \to \lp_0$ weakly as $N \to \infty$
    \comment{and they are conditionally Poisson},
    for any $T$ and $\delta > 0$ there exists a sequence
    $(C_1, C_2, \ldots)$ of finite constants such that
    \begin{align} \label{eqn:compact eqn lookdown}
    \limsup_{N \to \infty}
    \IP\left\{
        \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > C_n
        \text{ for some } n \ge 1
    \right\}
    < \delta .
    \end{align}
\end{lemma}

@@

\begin{proof}[of Lemma~\ref{lem:compact containment lookdown}]
    By a union bound,
    $$ 
    \IP\left\{
        \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > C_n
        \text{ for some } n \ge 1
    \right\}
    \le
    \sum_{n \ge 1}
    \IP\left\{
        \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > C_n
    \right\} ,
    $$
    so \eqref{eqn:compact eqn lookdown} follows
    by taking $C_n = 2^{n-1} K_n / \delta$ and using Lemma~\ref{lem:compact containment hn}.
\end{proof}

\begin{proof}[of Lemma~\ref{lem:compact containment hn}]
    \comment{TODO: just give bounds and say this follows
        the proof of Lemma~\ref{lem:eta_compact_containment} exactly, otherwise.
        Or make the basic idea a lemma in the appendix?}
    Let $h_n \in \mathcal{H}$.
    It is evident from the description of the process
    (or, by differentiating in Definition~\ref{defn:lookdown_mgale})
    that
    Applied to $f(x, u) = h_n(u)$, the martinagle representation~\eqref{eqn:f_xi_mgale} is
    \begin{align*}
    \begin{split}
        \langle h_n, \lp_t^N \rangle
        &=
        \langle h_n, \lp_0^N \rangle
        + M^{h_n}_t 
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            2 a^N(x, \eta^N_s) \int_u^N h_n(u_1) du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            h_n'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds ,
    \end{split}
    \end{align*}
    where $M^{h_n}_t$ is a martingale
    with quadratic variation
    \begin{align*}
        [M^{h_n}]_t
        &=
        \int_0^t
        \langle 2 a^N(x, \eta_s^N) \int_u^N h_n(u_1)^2 du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}
    Now, note that $0 \le a^N(x, \eta_x^N) \le C_a$ for some $C_a < \infty$,
    and we have assumed that $\int_u^\infty h_n(u_1) du_1 \le k_n h_n(u)$,
    so that 
    \begin{align*}
        & \int_0^t
        \big\langle
            2 a^N(x, \eta^N_s) \int_u^N h_n(u_1) du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad
        \le
        2 C_a k_n \int_0^t \big\langle h_n(u), \lp_s^N(dx, du) \big \rangle .
    \end{align*}
    In the next line, the contribution from $a^N(x, \eta^N_s) u^2 h'(u)$ is nonpositive because
    $a^N \ge 0$ and $h'(u) \le 0$.
    Similarly, since $b^N(x, \eta_x^N) \le C_b$ for some $C_b < \infty$,
    and by assumption
    $- h_n'(u) \le k_n h_n(u)$ and $h_n'(u) = 0$ for $u \ge u_n$,
    and so
    \begin{align*}
        & \int_0^t
        \big\langle
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            h_n'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \le
        C_b u_n k_n \int_0^t \big\langle h_n(u), \lp_s^N(dx, du) \big \rangle .
    \end{align*}
    Putting this together, we have that
    \begin{align} \label{eqn:hn_bound}
        \langle h_n, \lp_t^N \rangle
        &\le
        \langle h_n, \lp_0^N \rangle
        + M^{h_n}_t 
        + (2 C_a + C_b u_n) k_n \int_0^t 
            \langle h_n, \lp_s^N \rangle
        ds .
    \end{align}
    By Gronwall's inequality, with $C = (2 C_a + C_b u_n) k_n$
    \begin{align*}
        \IE\left[ \langle h_n, \lp_t^N \rangle \right]
        &\le
        \IE\left[ \langle h_n, \lp_0^N \rangle \right] e^{Ct} .
    \end{align*}

    Now consider the maximum. By \eqref{eqn:hn_bound},
    \begin{align*}
        \sup_{0 \le t \le T} \langle h_n, \lp_t^N \rangle
        &\le
        \langle h_n, \lp_0^N \rangle
        + \sup_{0 \le t \le T} M^{h_n}_t 
        + C \int_0^t 
            \sup_{0 \le t \le T} \langle h_n, \lp_s^N \rangle
        ds .
    \end{align*}
    Again since $\sqrt{x} \le 1 + x$ for $x \ge 0$,
    The Burkholder-Davis-Gundy inequality
    tells us that there is a $C'$ such that
    \begin{align*}
    \IE\left[ \sup_{0 \le t \le T} M^{h_n}_t \right]
    & \le
        C'\left( 1 + \IE\left[ [M^{h_n}]_t \right] \right)
    \\ & \le
        C' \left( 1 + \int_0^t \IE\left[
        \langle 2 a^N(x, \eta_s^N) \int_u^N h_n(u_1)^2 du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        \right] ds \right) 
    \\ & \le
        C' \left( 1 + 2 C_a k_n \int_0^t \IE\left[
        \langle h_n, \lp_s^N \big \rangle
        \right] ds \right) 
    \\ & \le 
        C' \left( 1 + C'' e^{Ct} \IE[ \langle h_n, \lp_t^N \rangle ] \right) ,
    \end{align*}
    where $C'' = 2 C_a k_n / C$.
    Therefore, 
    \begin{align*}
        \IE\left[ \sup_{0 \le t \le T} \langle h_n, \lp_t^N \rangle \right]
        &\le
        C' + \IE\left[ \langle h_n, \lp_0^N \rangle \right]( 1 + C'' e^{Ct} )
        + C \int_0^t 
            \IE\left[ \sup_{0 \le t \le T} \langle h_n, \lp_s^N \rangle \right]
        ds ,
    \end{align*}
    and so once again by Gronwall's inequality, there is a $C'''$ such that
    \begin{align*}
        \IE\left[ \sup_{0 \le t \le T} \langle h_n, \lp_t^N \rangle \right]
        &\le
        C'''(1 + \IE[\langle h_n, \lp_0^N \rangle]) e^{2Ct} ,
    \end{align*}
    and so
    \begin{align*}
        \IP\left\{ \sup_{0 \le t \le T} \langle h_n, \lp_t^N \rangle > K \right\}
        &\le
        \frac{
            C'''(1 + \IE[\langle h_n, \lp_0^N \rangle]) e^{2CT} 
        }{ K } .
    \end{align*}
\end{proof}

@@

\begin{proof}[of Lemma \ref{lem:test_fn_tightness}]
    Again, we use the Aldous-Rebolledo criterion.
    Tightness of $\langle f, \lp_t \rangle$ for a fixed $t$
    follows from Lemma~\ref{lem:compact containment hn},
    so we need only prove conditions analogous to
    \eqref{eqn:eta_projections_goal1}
    and \eqref{eqn:eta_projections_goal2}
    applied to the martingale representation
    of equations~\eqref{eqn:f_xi_mgale} and~\eqref{eqn:f_xi_qv}.
    Indeed, the proof follows exactly as the proof of
    Lemma~\ref{lem:eta_projections_tightness},
    except that terms are bounded by constant multiples of $\langle h_n, \xi^N_s \rangle$
    instead of $\langle 1, \eta^N_s \rangle$,
    where $h_n$ is bounded below on $[0, u_*]$.
    \comment{(or we could say that $u_n \ge 2 u_*$)}
    There is only one term that might use explanation,
    which is
    \begin{align*}
        &
        \frac{\theta}{N} \int_{\IR^d} \int_u^N (f(y, u) - f(x, u)) du_1 q^m(x, dy, \eta^N_s)
        \\ &\qquad
        % = \frac{\theta}{N} (N - u) \int_{\IR^d} (f(y, u) - f(x, u)) q^m(x, dy, \eta^N_s) \\ &\qquad
        = \left(1 - \frac{u}{N}\right) \theta \int_{\IR^d}
            (f(y, u) - f(x, u)) q^m(x, dy, \eta^N_s) ,
    \end{align*}
    but this is now clearly bounded by a constant multiple of $h_n(u)$.
\end{proof}
